Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-01-12 WSSSPE3/report/wg_appendix_industry_interaction.tex
Create wg_appendix_industry_interaction.tex
Reformatted group members to itemized list to ensure they appear in the right place.
Former-commit-id: e4584de998d67127bdd327661163fca316f83c01
adding POC request to appendices
Former-commit-id: 42aa40a42eb82bf40df6e482de5fb48dc9085b8d
Nic's changes, plus a couple of small edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Training Working Group Discussion}
\label{sec:appendix_training}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Group Members}
{\small
\begin{longtable}{ll}
   name            &  affiliation
\\ name2           &  affiliation2
\end{longtable}
}

@@

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Industry Interaction Working Group Discussion}
\label{sec:appendix_industry_interaction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Nic Weber\footnote{email: \href{mailto:nmweber@uw.edu}{nmweber@uw.edu}} will serve as the point of contact for this working group.

\subsection{Group Members}

\begin{itemize}
\item Nic Weber -- University of Washington
\item Suresh Marru -- Indiana University
\item Jeffrey Carver -- University of Alabama
\item Davide DelVento -- NCAR/CISL
\item Steven Brandt -- Louisiana State University
\end{itemize}

@@

Our initial broad question was, "What makes for successful transitions of scientific software from academia to industry?"

There are a number of potential funding transitions that may occur:  
%
\begin{itemize}

\item A project could be \textbf{refunded}, and development or maintenance of
the software continue as planned.

\item A project might locate a \textbf{new source of funding} in which case the
software may be further developed or simply maintained as before.

\item The project could transition to a \textbf{community supported model}
whereby ownership, maintenance, and stewardship of the software become similar
to peer-production models in open-source (e.g., see
Howison~\cite{howison_sustaining_2015}).

\item The project could receive some form of industry sponsorship in which case
ownership of the intellectual property, licensing, maintenance activities,
hosting, etc.\ may change significantly.

\end{itemize}

We characterized each of these potential changes in funding as ``transition
pathways'' to sustainable software (see similar work by Geels and
Schot~\cite{Geels:2007}).

Our work at WSSSPE3 included the following activities (described in detail
below): (1) brainstorming goals for this type of research, (2) imagining
potential outcomes of completing a set of case studies on this topic, and (3)
generating a set of working definitions for some of the broad concepts we are
describing.

First, we discussed the \textbf{goals} of this research, attempting to answer the 
question \emph{What is the goal of doing research on transition pathways?}
A number of research questions arose:  Can we identify collaborations that have 
occurred and try to understand which were successful, which were unsuccessful, 
and what factors contributed to these successes/failures? Can we determine what 
each partner wants to get out of such a collaboration? For example, why would 
industry be interested in collaborating with academia? Or why would academia 
be interested in collaborating with industry? How could we design a study that 
focused on the impact of the software in undergoing this type of transition?

Next, we imagined \textbf{potential outcomes} of research on this topic, involving 
a set of case studies that look at successful and unsuccessful
transitions of researchers between academia and industry. This might address 
each of the transition types (described below). Successful transitions are
described as those that lead to either weak or strong sustainability (also
defined below). In addition, the results from this research might help create a 
generalizable framework that might allow for the study of different transition 
pathways (other than academia to industry).

Finally, we created some \textbf{general definitions} for these concepts; we 
characterize transitions in the following ways:
\begin{itemize}

\item Handoff model: academia initially writes the software, industry (for-profit 
or nonprofit) then takes over the project.

\item Co-Production Model: industry and academia interact throughout development
of the project.

\item Sponsorship Model: academia writes and maintains the software; 
industry contributes funding for the development\slash maintenance of software.
In this example, industry is also likely a user of the software.

\item Spinoff model: transition to a for-profit or non-profit company owned by or in
collaboration with original developers.

\end{itemize}

We characterized sustainability in the following ways:
\begin{itemize}

\item Weak Sustainability: Software continues to be accessible, useful, and
usable.

\item Strong Sustainability: Software meets criteria above, but is also able to
be reused for further innovation (i.e., issued non-restrictive open-source
license).

\end{itemize}
We refer readers to Becker et al.~\cite{Becker:2014} for an extended discussion of weak versus strong
sustainability.

@@

Identify projects that are collaborative, perhaps by reviewing funded projects from programs specifically geared towards industry academic collaborations.

Develop a systematic process for conducting case studies (what kind of data are being gathered about each case)

### 2016-01-12 WSSSPE3/report/wg_main_industry_interaction.tex
Nic's changes, plus a couple of small edits

\subsection{Transition Pathways to Sustainable Software: Industry \& Academic Collaboration}

@@

There are a number of potential funding transitions that may occur:  
%
\begin{itemize}

\item A project could be \textbf{refunded}, and development or maintenance of
the software continue as planned.

\item A project might locate a \textbf{new source of funding} in which case the
software may be further developed or simply maintained as before.

\item The project could transition to a \textbf{community supported model}
whereby ownership, maintenance, and stewardship of the software become similar
to peer-production models in open-source (e.g., see
Howison~\cite{howison_sustaining_2015}).

\item The project could receive some form of industry sponsorship in which case
ownership of the intellectual property, licensing, maintenance activities,
hosting, etc.\ may change significantly.

\end{itemize}

We characterize each of these potential changes in funding as ``transition
pathways'' to sustainable software (see similar work by Geels and
Schot~\cite{Geels:2007}).

@@

Although transitions are often studied under the broad umbrella of ``technology
transfer,'' we believe there are likely to be a number of different ways in which
a pathway from initial production to long-term maintenance and secure funding are
achieved. In short, industry sponsorship is an important aspect of sustaining
scientific software, but our current understanding of these transitions focuses
narrowly on commercial successes\slash failures.

@@

In looking at existing literature that addresses industry transitions, many
reports focus on benefits that accrue to the private sector, or to a government
that originally sponsored the research project. This literature does not address
the impact that these transitions have on the accessibility or usability of the
software, or the impact that these transitions have on the career of the
researchers involved.

Examples of the former scenario (benefits accruing to private sector) are as
follows:
\begin{itemize}

\item REF Impact Case Studies: \url{http://impact.ref.ac.uk/CaseStudies/}

\item Background of projects funded in the UK: \url{http://gtr.rcuk.ac.uk/}

\item Dowling Review from the UK: addresses complexity of work between these two
communities: \url{http://www.raeng.org.uk/policy/dowling-review}

\item Pathway to Impact - UK report: two pages of grant proposals are asked to
forecast what impact they might have (including environmental, academic, economic).

\end{itemize}

@@

Our work at WSSSPE3 included the following activities (described in detail
below): (1) brainstorming goals for this type of research, (2) imagining
potential outcomes of completing a set of case studies on this topic, and (3)
generating a set of working definitions for some of the broad concepts we are
describing.

\textbf{Goal}

\emph{What is the goal of doing research on transition pathways?}

Can we identify collaborations that have occurred and try to understand which were successful, which were unsuccessful, and what factors contributed to these successes/failures?

Can we determine what each partner wants to get out of such a collaboration?
Why would industry be interested in collaborating with academia? 
Why would academia be interested in collaborating with industry?

How could we design a study that focused on the impact of the software in
undergoing this type of transition?

\textbf{Potential outcomes}

Complete a set of case studies which look at successful and unsuccessful
transitions of researchers between academia and industry. This might address 
each of the transition types (described below). Successful transitions are
described as those that lead to either weak or strong sustainability (also
defined below).

Create a generalizable framework that might allow for the study of different
transition pathways (other than academia to industry).

\textbf{General Definitions}

We characterize transitions in the following ways:
\begin{itemize}

\item Handoff model: academia initially writes the software, industry (for-profit 
or nonprofit) then takes over the project.

\item Co-Production Model: industry and academia interact throughout development
of the project.

\item Sponsorship Model: academia writes and maintains the software; the
industry contributes funding for the development\slash maintenance of software.
In this example, industry is also likely a user of the software.

\item Spinoff model: transition to for- or non-profit company owned by or in
collaboration with original developers.

\end{itemize}

We characterize sustainability in the following ways:
\begin{itemize}

\item Weak Sustainability: Software continues to be accessible, useful, and
usable.

\item Strong Sustainability: Software meets criteria above, but is also able to
be reused for further innovation (i.e., issued non-restrictive open-source
license).

\end{itemize}
We refer readers to Becker et al.~\cite{Becker:2014} for extended discussion of weak versus strong
sustainability. For more detail on the group's discussion, see
Appendix~\ref{sec:appendix_industry_interaction}.

@@

@@

To find more information about this group, or join it, see \todo{where should
someone go who want to know more about this and perhaps wants to contribute?}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transition Pathways to Sustainable Software: Industry \& Academic Collaboration}
\label{sec:industry_interaction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

@@

Most previous research and discussion of industry and academic collaboration, sharing, and funding of research software has focused on the impact of such arrangements. Examples of these types of reports are as follows:

\begin{itemize}
\item REF Impact Case Studies: \url{http://impact.ref.ac.uk/CaseStudies/}
\item Background of projects funded in the UK: \url{http://gtr.rcuk.ac.uk/}
\item Dowling Review from the UK: addresses complexity of work between these two
communities: \url{http://www.raeng.org.uk/policy/dowling-review}
\item Pathway to Impact - UK report: two pages of grant proposals are asked to
forecast what impact they might have (including environmental, academic, economic).
\end{itemize}

@@

Although sustainability transitions are often studied under the broad umbrella of ``technology
transfer,'' we believe there are likely to be a number of different ways in which
a pathway from initial production to long-term maintenance and secure funding is
achieved. In short, industry sponsorship is an important aspect of sustaining
scientific software, but our current understanding of these transitions focuses
narrowly on commercial successes or failures of those collaborations.

In looking at existing literature that addresses industry transitions, many
reports (such as those listed above) focus on benefits that accrue to the private sector, or to a government
that originally sponsored the research project. This literature does not address
the impact that these transitions have on the accessibility or usability of the
software, or the impact that these transitions have on the career of the
researchers involved.

For more detail on the group's discussion, see
Appendix~\ref{sec:appendix_industry_interaction}.

@@

The broad goals that we would like to accomplish are:

\begin{enumerate}
\item To complete a set of case studies which look at successful and unsuccessful transitions between academic researchers and industry
\item To create a generalizable framework, which might allow for a broader study of different transition pathways (other than between academia and industry)
\end{enumerate}

The main plan for the group going forward is the creation of a white paper on the topic of sustainability transitions.

@@

In lieu of a landing page, the Transitions Pathways
group requests an email be sent to Nic Weber\footnote{email: 
\href{mailto:nmweber@uw.edu}{nmweber@uw.edu}} to find out more
about the group's efforts and how to participate.
